On the specific recursive structure of iterated parametricity

### 2025-11-16 paper/paper.tex
Layout
paper: NFC whitespace changes
paper: remove unused \ap macro
Corrections, especially on the final step

% Some abbreviations
\newcommand{\tl}{\ensuremath{\mathsf{tl}}}
\newcommand{\hd}{\ensuremath{\mathsf{hd}}}
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}\;}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

@@

Since $\restrf[n][p]$ is referring to $\framep[n][p]$ and $\framep[n-1][p]$, and $\cohf[n][p]$ is referring to $\restrf[n][p]$ and $\restrf[n-1][p]$, thus also to $\framep[n][p]$, $\framep[n-1][p]$ and $\framep[n-2][p]$, we built all of these 6 components at the same time. But equations were also needed to state the properties of $\framep[n][p]$, $\framep[n-1][p]$ and $\restrf[n][p]$ so that we were building by induction on $n+p$ (that is on $n$, using the convention of \citep{hr25}) the following block of information:

@@

In the current paper and associated formalisation, we replace the equational properties by definitions, building instead by induction a block of the form:

\begin{align*}
  \framep[n][p] & : \U[m] & \mbox{for all $p\leq n$}                                                                                      \\
  \painting[n][p] & : \framep[n][p] \rightarrow \U[m] & \mbox{for all $p\leq n$}                                                          \\
  \restrf[n][p] & : \Pi q \leq n+p.\, \framep[n+1][p] \rightarrow \framep[n][p] & \mbox{for all $p\leq n$}                         \\
  E & : \fullframe[m][n+1] \imp \U[m] \\
  \restrp[n][p] & : \Pi q \leq n+p.\, \Pi d:\framep[n+1][p].\,\\
                  & \;\;\;\;\painting[n+1][p](d) \rightarrow \painting[n][p](\restrf[n][p](d)) & \mbox{for all $p\leq n$} \\
  \cohf[n][p]   & : \Pi qr\epsilon\omega.\, \restrf[n][p][\epsilon][q] \circ \restrf[n+1][p][\omega][r] = \\
                  & \;\;\;\;\restrf[n][p][\omega][r] \circ \restrf[n+1][p][\epsilon][q+1] & \mbox{for all $p\leq n$}                   \\
  E' & : \fullframe[m][n+2] \imp \U[m] \\
  \cohp[n][p]     & \Pi qr\epsilon\omega.\,\Pi d:\framep[n+2][p].\,
                  & \;\;\;\;\restrp[n][p][q][\epsilon](\restrf[n+1][p][r][\omega](d)) \circ \restrp[n+1][p][r][\omega](d) = \\
                  & \;\;\;\;\restrp[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon](d)) \circ \restrp[n+1][p][q+1][\epsilon](d) & \mbox{for all $p\leq n$}                 \\
\end{align*}
where $\framep[n+2][p]$, $\framep[n+1][p]$ and $\restrf[n+1][p]$, as
well as $\painting[n+2][p]$, $\painting[n+1][p]$ and $\restrp[n+1][p]$
are now appropriate definitions.

@@

So far, this means that we are able to specify a dependent structure of the form:
\begin{align*}
  \framep[n][p] & : \U & \mbox{for all $p\leq n$}                                                                                      \\
  \painting[n][p] & : \framep[n][p] \rightarrow \U & \mbox{for all $p\leq n$}                                                          \\
  \restrf[n][p] & : \Pi q \leq n+p.\, \framep[n+1][p]
    \left(\begin{array}{l}
      \framep[n][0,...,p-1]   \\
      \painting[n][0,...,p-1] \\
      \restrf[n][0,...,p-1]   \\
    \end{array}\right)
  \rightarrow \framep[n][p] & \mbox{for all $p\leq n$}                         \\
\end{align*}
which can appropriately be reformulated using sequences for appropriate definitions of $\frametype[n][0,...,p]$, $\paintingtype[n][0,...,p]$ and $\restrftype[n][0,...,p]$, since the dependencies of $\restrf[n][p]$ into $\restrf[n][0,...,p-1]$ makes it a very dependent function~\citep{Hickey96}:
\begin{align*}
  \framep[n][0,...,p] & : \frametype[n][0,...,p]                                                                                       \\
  \painting[n][0,...,p] & : \paintingtype[n][0,...,p](\framep[n][0,...,p])                                                         \\
  \restrf[n][0,...,p] & : \restrftype[n][0,...,p]
    \left(\begin{array}{l}
      \framep[n][0,...,p-1]   \\
      \painting[n][0,...,p-1] \\
      \restrf[n][0,...,p-1]   \\
    \end{array}\right)\\
\end{align*}

In general, the types of sequences such as $\restrf[n][0,...,p]$ will be nested since components, such as $\restrf[n][p]$, may depend on previous components of the sequence. So, the type of a sequence of length $p$ will be expressed as a $p$-iterated $\Sigma$-type. In the case of $\framep$ and $\painting$, there will be however no dependency, so an ordinary product can be used. If $\framep[n][0,...,p]$ is such a sequence of length $p+1$, we write $\framep[n][0,...,p-1]$ for its immediate subsequence of length $p$ and $\framep[n][p]$ for its last component, and similarly for all such sequences.

@@

    \depstype[n][0,..,p-1] & \defeq &
    \left(\begin{array}{lll}
              \framep[n][0,...,p-1]   & : & \frametype[n][0,...,p-1]    \\
              \painting[n][0,...,p-1] & : & \paintingtype[n][0,...,p-1]
              (\framep[n][0,...,p-1])                                   \\
              \restrf[n][0,...,p-1]   & : & \restrftype[n][0,...,p-1]
                  \framep[n][0,...,p-1]   \\
                  \painting[n][0,...,p-1] \\
    \framep[n+1][0,...,(p+1)-1]                                                                    &
    (\deps[n][0,...,p-1]:\depstype[n][0,...,p-1]) & :      & \frametype[n+1][0,...,p]                             \\
    \framep[n+1][0,...,(0+1)-1]                                                                          &
    (\unitpoint)                                                                 & \defeq & (\unitpoint,unit) \\
    \framep[n+1][0,...,((p+1)+1)-1]                                                                  &
    (\deps)                                                               & \defeq &
              \framep[n+1][0,...,p]
              (\deps[n][0,...,p-1]), \\
              \Sigma d:\framep[n+1][0,...,p]
              (\deps[n][0,...,p-1]) \\
              \quad \layer[n][p]
              (\deps[n][0,...,p])
              (d) \\
    \restrftype[n][0,...,p-1]                                                              &
              \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
              \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
            \end{array}\right)          & :      & \U                                      \\
    \restrftype[n][0,...,0-1]                                                              &
    \restrftype[n][0,...,(p+1)-1]                                                          &
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
      \Sigma \restrf[n][0,...,p-1]
      :\restrftype[n][0,...,p-1]
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
      \Pi q\leq n.\,\Pi \omega.\,\framep[n+1][p]
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
                \restrf[n][0,...,p-1]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}                                                    \\
    \layertype[n][p]                                                                       &
    (\deps[n][0,...,p]:\depstype[n][0,...,p])
    & :      & \Type               \\
    \layertype[n][p]                                                                       &
    (\deps[n][0,...,p])
    & \defeq & \framep[n+1][p]
    (\deps[n][0,...,p-1])
    \rightarrow \U                                                                               \\
    \layer[n][p]                                                                           &
    (\deps[n][0,...,p]:\depstype[n][0,...,p])
    & :      & \layertype[n][p]
    (\deps[n][0,...,p])
    \layer[n][p]                                                                           &
    (\deps[n][0,...,p])
    d \mapsto \Pi\omega.\,\painting[n][p](\restrf[n][p](d))\\

@@

We can now formally define $\painting[n+1][p]$ by recursion on $n+1-p$:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][p]      &
    (\fulldeps[n]:\fulldepstype[n])                     \\
    \painting[n+1][n]      &
    (\fulldeps[n])     & \defeq & \fulldeps.E             \\
    \painting[n+1][p](p<n) &
    (\fulldeps[n])     & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\fulldeps.\deps[n][0,...,p]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                             \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][0,...,p]   &
    (\fulldeps[n-1])        & :      & \paintingtype[n+1][0,...,p](\framep[n+1][0,...,p](\fulldeps.\deps[n-1][0,...,p])) \\
    \painting[n+1][0,...,0-1] &
    E                       & \defeq & \unittype                                                                     \\
    \painting[n+1][0,...,p]   &
    (\fulldeps[n-1])
                            & \defeq &
    (\painting[n+1][0,...,p-1](\fulldeps[n-1]),\painting[n+1][p](\fulldeps[n-1]))                                        \\
  \end{array}
\end{equation*}

@@

    \depscohstype[n][0,..,p-1] & \defeq &

@@

    \overline{\deps}(\depscohs[n][0,...,p-1]:\depscohstype[n][0,...,p-1]) & :      & \depstype[n+1][0,...,p] \\
    \overline{\deps}(\depscohs)               & \defeq &

@@

    (E)

@@

      & :      & \U                                                                               \\

@@

      (\depscohs[n][0,..,p-1])

@@

    \end{array}\right.\\

@@

      (\depscohs[n][0,..,p]:\depscohstype[n][0,..,p]) & :      &
    \restrltype[n][p] (\depscohs[n][0,..,p])                                                   \\

@@

@@

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \fulldepstype[n+1] \\
  \overline{\fulldeps}(\fulldepscohs[n])               & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\fulldepscohs.\depscohs[n][0,...,n-1])\\
            \fulldepscohs.\depscohs.\fulldeps.E\\
          \end{array}\right)             \\
\end{align*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n+1][0,...,p]   &
    (\fulldepscohs[n])        & :      & \restrptype[n+1][0,...,p](\fulldepscohs.\depscohs.\fulldeps[n])) \\
    \restrp[n+1][0,...,0-1] &
    E  & \defeq & \unittype                                                                   \\
    \restrp[n+1][0,...,p]   &
    (\fulldepscohs[n])
                            & \defeq &
    (\restrp[n+1][0,...,p-1](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                                        \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p-1] & (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \U                                      \\
    \cohptype[n][0,...,0-1]        & \unitpoint                                        & \defeq & \unittype                               \\
    \cohptype[n][0,...,(p+1)-1] & (\fulldepscohs[n])                           & \defeq & \cohptype[n][0,...,p-1](\fulldepscohs[n]) \times \\
    & & & \;\;\;\;\left\{
    \begin{array}{l}
        \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d:\framep[n+2][p](\overline{\deps}(\fulldepscohs.\depscohs[n][0,...,p-1])).\,\Pi c:\painting[n+2][p](\overline{\fulldeps}(\fulldepscohs[n]))\\
        \fulldepscohs.\depscohs.\restrp[n][p][q][\epsilon](\restrf[n+1][p](d))(\restrp[n+1][p][r][\omega]
        (\fulldepscohs[n])(d)(c)) =                         \\
        \fulldepscohs.\depscohs.\restrp[n][p][r][\omega](\restrf[n+1][p](d))(\restrp[n+1][p][q+1][\epsilon]
        (\fulldepscohs[n])(d)(c))
      \end{array}
    \right.                                                                                                           \\
  \end{array}
\end{equation*}

@@

            \end{array}\right) & :      & \cohftype[n+1][0,...,p]                                                                                             \\

@@

            \end{array}\right)            & :      & \U                                                                                                        \\

@@

            \end{array}\right)                                  & :      &

@@

    \end{array}                                         \\

@@

We need also the notion of full dependencies at the level of the coherence conditions of dimension 2, that combines a full $\depscoht[n][0,...,n]$ with a set $E:\framep[n+3][n+2](\overline{\depscohs}(\depscoht[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohttype[n] \defeq
  \left(\begin{array}{lll}
            \depscoht[n][0,...,n-1] &:&\depscohttype[n][0,...,n-1]    \\
            E&:&\framep[n+3][n+2](\overline{\depscohs}(\depscoht[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}

@@

\begin{align*}
  \overline{\fulldepscohs}(\fulldepscoht:\fulldepscohstype) & :      & \fulldepscohstype \\
  \overline{\fulldepscohs}(\fulldepscoht)                   & \defeq &
  \left(\begin{array}{l}
            \overline{\depscohs}(\fulldepscoht.\depscoht) \\
            \overline{\depscohs}(\fulldepscoht.\depscoht[n][0,...,n-1])\\
            \fulldepscoht.\depscoht.\fulldepscohs.E\\
          \end{array}\right)             \\
\end{align*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n+1][0,...,p]   &
    (\fulldepscoht[n])        & :      & \cohptype[n+1][0,...,p](\fulldepscoht.\depscoht.\fulldepscohs[n])) \\
    \cohp[n+1][0,...,0-1] &
    E  & \defeq & \unittype                                                                   \\
    \cohp[n+1][0,...,p]   &
    (\fulldepscoht[n])
                            & \defeq &
    (\cohp[n+1][0,...,p-1](\fulldepscoht[n]),\cohp[n+1][p](\fulldepscoht[n]))                                        \\
  \end{array}
\end{equation*}

@@

Armed with all the constructions and assuming enough $E:\framep[n][n-1]
\rightarrow \U$ given, we can now define $\framep[n][p]$,
$\painting[n][p]$, $\restrf[n][p]$, $\restrp[n][p]$, $\cohf[n][p]$,
and $\cohp[n][p]$ for all $n$ and $p$. It is indeed enough to build
$\fulldepscoht[n]$ by induction on $n$:
\begin{itemize}
\item $\fulldepscohs[n+1]$ is $\overline{\fulldepscohs}(\fulldepscoht[n])$
\item $\cohp[n+1][0,...,n]$ is given from $\fulldepscoht$ as above
\item $\coht[n+1][0,...,n]$ is given by uniqueness of identity proofs
\end{itemize}

>>>>

% Some abbreviations
\newcommand{\tl}{\ensuremath{\mathsf{tl}}}
\newcommand{\hd}{\ensuremath{\mathsf{hd}}}
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

@@

Since $\restrf[n][p]$ is referring to $\framep[n][p]$ and $\framep[n-1][p]$, and $\cohf[n][p]$ is referring to $\restrf[n][p]$ and $\restrf[n-1][p]$, thus also to $\framep[n][p]$, $\framep[n-1][p]$ and $\framep[n-2][p]$, we built all of these 6 components at the same time. But equations were also needed to state the properties of $\framep[n][p]$, $\framep[n-1][p]$ and $\restrf[n][p]$ so that we were building by induction on $n+p$ (that is on $n$, using the convention of \cite{hr25}), for any $D:\Xto{m}{n}$, the following block of information:

@@

In the current paper and associated formalisation, we switch the point
of view, take $\framep[n-2][p]$ $\painting[n-2][p]$, $\restrf[n-1][p]$
and $\restrp[n-1][p]$ as references and renaming them respectively as
$\framep[n][p]$, $\painting[n][p]$, $\restrf[n][p]$ and
$\restrp[n][p]$, replace the equational properties of $\framep[n][p]$,
$\framep[n-1][p]$, $\restrf[n][p]$, $\painting[n][p]$,
$\painting[n-1][p]$ and $\restrp[n][p]$ as definitions of objects
respectively renamed to $\framep[n+2][p]$, $\framep[n+1][p]$,
$\restrf[n+1][p]$, $\painting[n+2][p]$, $\painting[n+1][p]$ and
$\restrp[n+1][p]$, building instead by induction a block of the form:

\begin{align*}
  \framep[n][p]   & : \U[m]                                                                                                     & \mbox{for all $p\leq n$} \\
  \painting[n][p] & : \framep[n][p] \rightarrow \U[m]                                                                           & \mbox{for all $p\leq n$} \\
  \restrf[n][p]   & : \Pi q \leq n+p.\, \Pi \epsilon.\, \framep[n+1][p] \rightarrow \framep[n][p]                                               & \mbox{for all $p\leq n$} \\
  \restrp[n][p]   & : \Pi q \leq n+p.\, \Pi \epsilon.\, \Pi d:\framep[n+1][p].\,                                                                                           \\
                  & \;\;\;\;\painting[n+1][p](d) \rightarrow \painting[n][p](\restrf[n][p](d))                                  & \mbox{for all $p\leq n$} \\
  \cohf[n][p]     & : \Pi qr\epsilon\omega.\, \restrf[n][p][\epsilon][q] \circ \restrf[n+1][p][\omega][r] =                                                \\
                  & \;\;\;\;\restrf[n][p][\omega][r] \circ \restrf[n+1][p][\epsilon][q+1]                                       & \mbox{for all $p\leq n$} \\
  \cohp[n][p]     & \Pi qr\epsilon\omega.\,\Pi d:\framep[n+2][p].\,\\
                  & \;\;\;\;\restrp[n][p][q][\epsilon](\restrf[n+1][p][r][\omega](d)) \circ \restrp[n+1][p][r][\omega](d) =                                \\
                  & \;\;\;\;\restrp[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon](d)) \circ \restrp[n+1][p][q+1][\epsilon](d) & \mbox{for all $p\leq n$} \\
\end{align*}

@@

So far, using this recursive definition of  $\framep[n][0,...,p-1]$ and $\restrf[n][0,...,p-1]$, we are able to specify the three first component of the exected dependent structure:
\begin{align*}
  \framep[n][p]             & : \U                                & \mbox{for all $p\leq n$} \\
  \painting[n][p]           & : \framep[n][p] \rightarrow \U      & \mbox{for all $p\leq n$} \\
  \restrf[n][p]             & : \Pi q \leq n+p.\, \framep[n+1][p]
  \left(\begin{array}{l}
            \framep[n][0,...,p-1]   \\
            \painting[n][0,...,p-1] \\
            \restrf[n][0,...,p-1]   \\
          \end{array}\right)
  \rightarrow \framep[n][p] & \mbox{for all $p\leq n$}                                       \\
\end{align*}
which can appropriately be reformulated using sequences for appropriate definitions of $\frametype[n][0,...,p]$, $\paintingtype[n][0,...,p]$ and $\restrftype[n][0,...,p]$, since the dependencies of $\restrf[n][p]$ into $\restrf[n][0,...,p-1]$ makes it a very dependent function~\citep{Hickey96}:
\begin{align*}
  \framep[n][0,...,p]   & : \frametype[n][0,...,p]                         \\
  \painting[n][0,...,p] & : \paintingtype[n][0,...,p](\framep[n][0,...,p]) \\
  \restrf[n][0,...,p]   & : \restrftype[n][0,...,p]
  \left(\begin{array}{l}
            \framep[n][0,...,p-1]   \\
            \painting[n][0,...,p-1] \\
            \restrf[n][0,...,p-1]   \\
          \end{array}\right)                                            \\
\end{align*}

The type of sequences such as $\restrf[n][0,...,p]$ will be represented by iterated $\Sigma$-types since each $\restrf[n][p]$ may depend on the previous components of the sequence. In the case of $\framep$ and $\painting$, there will be however no dependency, so an ordinary product can be used to represent $\framep[n][0,...,p]$ and $\painting[n][0,...,p]$. Later on, we will write $\framep[n][0,...,p-1]$ for the immediate subsequence of length $p$ of $\framep[n][0,...,p]$ and $\framep[n][p]$ for its last component, and similarly for $\painting[n][0,...,p]$, $\restrf[n][0,...,p]$, etc.

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \depstype[n][0,..,p-1]          &        & \defeq            &
    \left(\begin{array}{lll}
              \framep[n][0,...,p-1]   & : & \frametype[n][0,...,p-1]    \\
              \painting[n][0,...,p-1] & : & \paintingtype[n][0,...,p-1]
              (\framep[n][0,...,p-1])                                   \\
              \restrf[n][0,...,p-1]   & : & \restrftype[n][0,...,p-1]
              \left(\begin{array}{l}
                  \framep[n][0,...,p-1]   \\
                  \painting[n][0,...,p-1] \\
                \end{array}\right)                             \\
            \end{array}\right)        \\
    \framep[n+1][0,...,(p+1)-1]     &
    \multicolumn{3}{l}{
    (\deps[n][0,...,p-1]:\depstype[n][0,...,p-1]) :\frametype[n+1][0,...,p]}         \\
    \framep[n+1][0,...,(0+1)-1]     &
    (\unitpoint)                    & \defeq & (\unitpoint,unit)                     \\
    \framep[n+1][0,...,((p+1)+1)-1] &
    \deps[n][0,...,p]               & \defeq &
    \left(\begin{array}{l}
              \framep[n+1][0,...,p]
              (\deps[n][0,...,p-1]), \\
              \Sigma d:\framep[n+1][0,...,p]
              (\deps[n][0,...,p-1])  \\
              \quad \layer[n][p]
              (\deps[n][0,...,p])
              (d)                    \\
            \end{array}\right)                                              \\
    \restrftype[n][0,...,p-1]       &
    \multicolumn{3}{l}{
    \left(\begin{array}{l}
              \framep[n][0,...,p-1]:\frametype[n][0,...,p-1]                             \\
              \painting[n][0,...,p-1]:\paintingtype[n][0,...,p-1](\framep[n][0,...,p-1]) \\
            \end{array}\right) : \U} \\
    \restrftype[n][0,...,0-1]       &
    \left(\begin{array}{l}
              \unitpoint \\
              \unitpoint \\
            \end{array}\right)          & \defeq &
    \unittype                                                                        \\
    \restrftype[n][0,...,(p+1)-1]   &
    \left(\begin{array}{l}
              \framep[n][0,...,p]   \\
              \painting[n][0,...,p] \\
            \end{array}\right)        & \defeq &
    \begin{array}{l}
      \Sigma \restrf[n][0,...,p-1]
      :\restrftype[n][0,...,p-1]
      \left(\begin{array}{l}
                \framep[n+1][0,...,p-1]   \\
                \painting[n+1][0,...,p-1] \\
              \end{array}\right).                     \\
      \Pi q\leq n.\,\Pi \omega.\,\framep[n+1][p]
      \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
                \restrf[n][0,...,p-1]   \\
              \end{array}\right) \rightarrow  \framep[n][p] \\
    \end{array}               \\
    \layertype[n][p]                &
    \multicolumn{3}{l}{
      (\deps[n][0,...,p]:\depstype[n][0,...,p])
    : \U}                                                                            \\
    \layertype[n][p]                &
    \deps[n][0,...,p]
                                    & \defeq & \framep[n+1][p]
    (\deps[n][0,...,p-1])
    \rightarrow \U                                                                   \\
    \layer[n][p]                    &
    \multicolumn{3}{l}{
      (\deps[n][0,...,p]:\depstype[n][0,...,p])
      : \layertype[n][p]
      (\deps[n][0,...,p])}
    \\
    \layer[n][p]                    &
    \deps[n][0,...,p]
                                    & \defeq &
    d \mapsto \Pi\omega.\,\painting[n][p](\restrf[n][p](d))                          \\
  \end{array}
\end{equation*}

@@

We can now formally define $\painting[n+1][p]$ by recursion on $n+1-p$:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][p]      &
    (\fulldeps[n]:\fulldepstype[n])                           \\
    \painting[n+1][n]      &
    \fulldeps[n]         & \defeq & \fulldeps.E             \\
    \painting[n+1][p](p<n) &
    \fulldeps[n]         & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer[n][p](\fulldeps.\deps[n][0,...,p]). \\
              \painting[n+1][p+1]
              (\fulldeps[n])(d,l)                                 \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n+1][0,...,p]   &
    (\fulldeps[n]:\fulldepstype[n])          & :      & \paintingtype[n+1][0,...,p](\framep[n+1][0,...,p](\fulldeps.\deps[n-1][0,...,p])) \\
    \painting[n+1][0,...,0-1] &
    \fulldeps[n]                         & \defeq & \unittype                                                                         \\
    \painting[n+1][0,...,p]   &
    \fulldeps[n]
                              & \defeq &
    (\painting[n+1][0,...,p-1](\fulldeps[n-1]),\painting[n+1][p](\fulldeps[n]))                                          \\
  \end{array}
\end{equation*}

@@

    \depscohstype[n][0,..,p-1]      &                         & \defeq                                                         &

@@

    \overline{\deps}                &
    (\depscohs[n][0,...,p-1]:\depscohstype[n][0,...,p-1]) : \depstype[n+1][0,...,p]}                                                                                                                             \\
    \overline{\deps}                & \depscohs[n][0,...,p-1] & \defeq                                                         &

@@

    E

@@

    : \U}                                                                                                                                                                                                        \\

@@

    \depscohs[n][0,..,p-1]

@@

@@

      (\depscohs[n][0,..,p]:\depscohstype[n][0,..,p]) :
    \restrltype[n][p] (\depscohs[n][0,..,p])}                                                                                                                                                                    \\

@@

    \depscohs[n][0,..,p]            & \defeq                  &

@@

We can then extend $\overline{\deps}$ to build $\fulldeps$ from $\fulldepscohs$:
\begin{align*}
  \overline{\fulldeps}(\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \fulldepstype[n+1] \\
  \overline{\fulldeps}(\fulldepscohs[n])                      & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\fulldepscohs.\depscohs[n][0,...,n-1]) \\
            \fulldepscohs.\depscohs.\fulldeps.E                     \\
          \end{array}\right)                             \\
\end{align*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n+1][0,...,p]   &
    (\fulldepscohs[n]:\fulldepscohstype[n])      & :      & \restrptype[n+1][0,...,p](\fulldepscohs.\depscohs.\fulldeps[n])) \\
    \restrp[n+1][0,...,0-1] &
    \fulldepscohs[n]                       & \defeq & \unittype                                                        \\
    \restrp[n+1][0,...,p]   &
    \fulldepscohs[n]
                            & \defeq &
    (\restrp[n+1][0,...,p-1](\fulldepscohs[n]),\restrp[n+1][p](\fulldepscohs[n]))                       \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p-1]     & (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \U                                               \\
    \cohptype[n][0,...,0-1]     & \unitpoint                              & \defeq & \unittype                                        \\
    \cohptype[n][0,...,(p+1)-1] & (\fulldepscohs[n])                      & \defeq & \cohptype[n][0,...,p-1](\fulldepscohs[n]) \times \\
                                &                                         &        & \;\;\;\;\left\{
    \begin{array}{l}
      \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d:\framep[n+2][p](\overline{\deps}(\fulldepscohs.\depscohs[n][0,...,p-1])).\,\Pi c:\painting[n+2][p](\overline{\fulldeps}(\fulldepscohs[n])) \\
      \fulldepscohs.\depscohs.\restrp[n][p][q][\epsilon](\restrf[n+1][p](d))(\restrp[n+1][p][r][\omega]
      (\fulldepscohs[n])(d)(c)) =                                                                                                                                                                              \\
      \fulldepscohs.\depscohs.\restrp[n][p][r][\omega](\restrf[n+1][p](d))(\restrp[n+1][p][q+1][\epsilon]
      (\fulldepscohs[n])(d)(c))
    \end{array}
    \right.                                                                                                                           \\
  \end{array}
\end{equation*}

@@

              \end{array}\right) : \cohftype[n+1][0,...,p]}                                                                \\

@@

            \end{array}\right)      : \U}                                                                \\

@@

            \end{array}\right)                                 :

@@

      \end{array}} \\

@@

We need also the notion of full dependencies at the level of the coherence conditions of dimension 2, that combines a full $\depscoht[n][0,...,n]$ with a set $E:\framep[n+3][n+2](\overline{\depscohs}(\depscoht[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohttype[n] \defeq
  \left(\begin{array}{lll}
            \depscoht[n][0,...,n-1] & : & \depscohttype[n][0,...,n-1]                                                    \\
            E                       & : & \framep[n+3][n+2](\overline{\depscohs}(\depscoht[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}

@@

\begin{align*}
  \overline{\fulldepscohs}(\fulldepscoht:\fulldepscohstype) & :      & \fulldepscohstype \\
  \overline{\fulldepscohs}(\fulldepscoht)                   & \defeq &
  \left(\begin{array}{l}
            \overline{\depscohs}(\fulldepscoht.\depscoht)               \\
            \overline{\depscohs}(\fulldepscoht.\depscoht[n][0,...,n-1]) \\
            \fulldepscoht.\depscoht.\fulldepscohs.E                     \\
          \end{array}\right)                      \\
\end{align*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohp[n+1][0,...,p]   &
    (\fulldepscoht[n]:\fulldepscohttype[n])    & :      & \cohptype[n+1][0,...,p](\fulldepscoht.\depscoht.\fulldepscohs[n])) \\
    \cohp[n+1][0,...,0-1] &
    \fulldepscohs[n]      & \defeq & \unittype                                                          \\
    \cohp[n+1][0,...,p]   &
    \fulldepscoht[n]
                          & \defeq &
    (\cohp[n+1][0,...,p-1](\fulldepscoht[n]),\cohp[n+1][p](\fulldepscoht[n]))                           \\
  \end{array}
\end{equation*}

@@

Armed with all the constructions, we are now able to specify the whole
structure on which we can reason by induction, for $D:\Xto{m}{n}$
given:

\begin{align*}
  \framep[n][0,...,n-1]   & : \frametype[n][0,...,n-1]                         \\
  \painting[n][0,...,n-1] & : \paintingtype[n][0,...,n-1](\framep[n][0,...,n-1]) \\
  \restrf[n][0,...,n-1]   & : \restrftype[n][0,...,n-1]
    \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
    \end{array}\right)                                            \\
  \restrp[n][0,...,n-1](E:\framep[n+1][n]
    \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
    \end{array}\right)\rightarrow \U) & : \restrptype[n][0,...,n-1]
    \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
              E\\
    \end{array}\right)                                            \\
  \cohf[n][0,...,n-1](E:\framep[n+1][n]
    \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
    \end{array}\right)\rightarrow \U) & : \cohftype[n][0,...,n-1]
    \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
              E\\
              \restrp[n][0,...,n-1]   \\
    \end{array}\right)                                            \\
    \cohp[n][0,...,n-1]
      \left(\begin{array}{l}
              E:\framep[n+1][n]
                \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
        \end{array}\right)\rightarrow \U \\
              E':\framep[n+2][n+1](\overline{\deps}
              \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
              E\\
              \restrp[n][0,...,n-1]   \\
              \cohf[n][0,...,n-1]   \\
        \end{array}\right))\rightarrow \U
      \end{array}\right)                                            \\
      & : \cohptype[n][0,...,n-1]
      \left(\begin{array}{l}
              \framep[n][0,...,n-1]   \\
              \painting[n][0,...,n-1] \\
              \restrf[n][0,...,n-1]   \\
              E\\
              \restrp[n][0,...,n-1]   \\
              \cohf[n][0,...,n-1]   \\
              E'\\
      \end{array}\right)                                            \\
\end{align*}
